---
title: "For Live Session - Unit 3"
author: "Hudson"
date: "1/15/2022"
output: html_document
---

#### Part 1: Part 1: We would like to analyze the Left Midfielders (LM) versus the Left Forwards (LF). (Estimated / expected time: 2 – 4 hours and at least 2+ slides).###

Part 1.1: Using the FIFA player data set, filter the data set to create a dataframe that has just the Left Midfielders (LM) and Left Forwards (LF).  
```{r}
#load the FIFA_Players data into fifa object for ease in future coding 
fifa = FIFA_Players

###filter the data set to create a dataframe that has just the Left Midfielders (LM) and Left Forwards (LF).
pos_lm_lf <- fifa %>% filter(Position == "LM" | Position == "LF")

```


### Part 1.2:Use Ggally and ggpairs() and the dataset you created above above, to plot the categorical variable Position (LM and LF), versus the continuous variables Acceleration and Agility.  
```{r}
#load the FIFA_Players data into fifa object for ease in future coding
fifa = FIFA_Players
library(GGally)

#plot the categorical variable Position (LM and LF), versus the continuous variables Acceleration and Agility.
pos_lm_lf %>% select(Acceleration, Agility, Position) %>% ggpairs(aes(color = Position),lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1)))

```


#Part 1.3: Given the plot above, what relationships do you see?  Comment on these.
```{r}
#Zooming in the Agility vs Acceleration plot from the ggpairs() plot in part 1.2, there appears to be a position relationship between agility and acceleration. That is, as the player’s acceleration increases, the players’s agility naturally will increase. This makes intuitive sense as agility is dependent on the player’s ability to accelerate when stopping/starting or changing direction.  


```

#Part 1.4: Your client would like to formally test if the mean agility rating of left midfielders is different than that of the left forwards.  Perform a 6 – step t-test to test for the difference in these means.(You may skip step 2 (draw and shade) if you like.  If you are unfamiliar with the 6-step hypothesis test, see Stat 1 slides or the Bridge Course to review the 6-step hypothesis test.)
```{r}
#load the FIFA_Players data into fifa object for ease in future coding
fifa <- FIFA_Players
library(GGally)

#load get the averages of the agility ratings for each of the positions (LM) 
LM_ag <- fifa %>% filter(Position == "LM") %>% filter(!is.na(Agility)) %>% select(Agility) #%>% arrange()
hist(LM_ag$Agility, main = "Histogram of LM Agility Ratings")

#test the normality of the agility ratings for each of the positions (LM)
qqnorm(LM_ag$Agility, main = 'LM Normal')
qqline(LM_ag$Agility)

#load get the averages of the agility ratings for each of the positions (LF) 
LF_ag <- fifa %>% filter(Position == "LF") %>% filter(!is.na(Agility)) %>% select(Agility) #%>% arrange()
hist(LF_ag$Agility, main = "Histogram of LF Agility Ratings")

#test the normality of the agility ratings for each of the positions (LF)
qqnorm(LF_ag$Agility, main = 'LF Normal')
qqline(LF_ag$Agility)

#load get the averages of the agility ratings for each of the positions (LM & LF) 
ag <- fifa %>% filter(Position == "LM" | Position == "LF")%>% filter(!is.na(Agility)) %>% select(Agility) 
hist(ag$Agility, main = "Histogram of Agility Ratings (Both Groups)")

#test the normality of the agility ratings for each of the positions (LF & LF)
qqnorm(ag$Agility, main = 'Normal')
qqline(ag$Agility)

#print the averages of the agility ratings for each of the positions (LM & LF) fir sanity check
LM_ag
LF_ag

#load the two sample t-test with the assumption of unequal varience between LF agility and LM agility
ttest <- t.test(LF_ag, LM_ag, alternative = "two.sided", var.equal = FALSE)
ttest
```


#Part 1.5:Are the assumptions of this test reasonably met?  If you have not had Stat 1, simply create a histogram of the agility scores for both groups (LM and LF) and given what you know about the CLT, comment on if you believe the sampling distribution of sample means (of your sample size) will be reasonably normal.  In addition, does there look like there is significant visual evidence to suggest the standard deviations are different? ….. If you have had Stat 1, create the plots listed above (and any other plots you might prefer) and be prepared to be a teacher and teach what you know about the assumptions of the t-test and if those are assumption are reasonably met here.
```{r}
#load the FIFA_Players data into fifa object for ease in future coding
fifa <- FIFA_Players
library(GGally)

#load get the averages of the agility ratings for each of the position LF
LF_ag <- fifa %>% filter(Position == "LF") %>% filter(!is.na(Agility)) %>% select(Agility, Position) 

#store the population Agility datafreame for LM positions 
LF <- data.frame(LF_ag)

#test for normality on the Agility dataframe for LF positions
qqnorm(LF$Agility, main = 'LF Normal')
qqline(LF$Agility)


#CLT for the larger sample 'LM'
n = 15 # number of samples to align with population of 'LF' data
iter= 15 # number of iterations = 1095 data points

Agility_CLT = numeric(iter)
Pos <- fifa %>% filter(Position == "LM") %>% filter(!is.na(Agility)) %>% select(Agility, Position)
Position <- Pos
#load the samples for CLT
for(i in 1:iter){
  samp = sample_n(LM_ag, n)
  xbar = mean(samp$Agility)
  Agility_CLT[i] = xbar
}

#convert double values to numeric to allow for data type pairing later on 
Agility_i <- as.integer(Agility_CLT)
Agility_s <-as.character(Agility_i)
Agility <- as.numeric(Agility_s)

#construct the CLT Agility dataframe for LM positions
CLT_ag_lm <- data.frame(Agility) 
Agility <- CLT_ag_lm
Position <- c("LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM", "LM")
CLT_LM <- data.frame(CLT_ag_lm, Position)

#test for normality on the CLT Agility dataframe for LM positions
qqnorm(CLT_LM$Agility, main = 'CLT LM Normal')
qqline(CLT_LM$Agility)


#pair up the the sample CLT Agility Dataframe for LM positions and population Agility dataframe for LF positions
ag_comp <- rbind(LF, CLT_LM)


#ggplot the faceted data for both sample CLT Agility Dataframe for LM positions and population Agility dataframe for LF positions
ag_comp %>% ggplot() + geom_histogram(mapping = aes(x = Agility, fill = Position), binwidth = 2)+ ggtitle("Agility vs Position") + xlab("Agility") + ylab("Count of Players")+ coord_cartesian(xlim = c(50, 100), ylim = c(0,10))+facet_wrap(~Position)


```

### Part 2: Part 2: (Estimated / expected time 3-5 hours and at least 3+ slides) 
Select/create at least 2 categorical variables and select two continuous variables and perform an EDA.  Also, at least one of the categorical variables should be created from a continuous variable (using the cut() function). ###   

#Part 2.1:Use these variables to explore the data and tell a story of what you discovered similar to what was shown in the asynch videos.  You do not need to go so far as to use linear regression, but let your curiosity guide you along the way and feel free to use methods you are familiar with that are appropriate to answering those questions. Your evidence could be purely visual or could include additional methods, it is up to you… just do your best and have fun!
```{r}
#load the FIFA_Players data into fifa object for ease in future coding
fifa <- FIFA_Players
library(GGally)

#make age category from fifa
fifa_ar <- fifa%>% filter(!is.na(Agility), !is.na(Overall), !is.na(Age), !is.na(PreferredFoot)) %>% mutate(AgeRange =  cut(Age, breaks = c(0, 20, 25, 30, 35, 40, 100), labels = c("<20", "21-25", "26-30", "31-35", "36-40", ">40")), .after = Age)
fifa_ar

#plot the categorical variables Preferred Foot, AgeRange versus the continuous variables Overall and Agility.
fifa_ar %>% select(Agility, Overall, PreferredFoot, AgeRange) %>% ggpairs(aes(color = AgeRange),lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1)))

```


