---
title: "For Live Session Unit 5"
author: "Hudson"
date: "1/27/2022"
output: html_document
---

Part 1: BALL STUDY
We previously used a dataset called PlayerBBall.csv which contained information about NBA basketball players.  To finish that assignment, you had to manipulate the height column.  Review the code you used to do that and see if you can’t make more efficient code using regular expressions and / or the string functions from this Unit. 
```{r}
#load in the needed packages for this code
library(stringr)
library(tidyft)
library(ggplot2)
library(tidyr)
library(readr)
library(plotly)

#ensure prints all values in dataframe
options(max.print = 1000000)

#load in the PlayersBBall.csv data
pd <- PlayersBBall

#part 1.1
#Use regular expressions to use the height column to create a TotalInches column that has the total height in inches and is recorded as a numeric variable. 

#Create numeric vector based on "height" 
pdh <- data.frame(x = pd$height)

#separate the height values to make numeric 
pdh_in <- pdh%>%separate(x,c("ft", "inch"), sep = "-")
ft <- as.numeric(pdh_in$ft)
inch <- as.numeric(pdh_in$inch)

#convert ft column units from feet to inches 
ft2in <- ft*12

#add the two column values together and store in TotalInches
total_inches <- ft2in+inch

#part 1.2
#Use this variable to make a chart that contains histograms of heights for every position (color coded).  

#prepare the position for histogram
Positions <- pd$position 

#Creating dataframe for just the variables i need for the histogram
TotalInches <- data.frame(total_inches, Positions)

#Plot the players height in inches on a hisogram and color fill the data for positions
TotalInches %>% ggplot(mapping = aes(x = total_inches, fill = Positions))+ geom_histogram()+xlab("Height [in]")+ylab("Player Count")+ggtitle("Histogram: Player Height")

```

Part 2: FIFA STUDY
We previously used a dataset called FIFA Players.csv which contained information about Soccer players.  
```{r}
#load in the needed packages for this code
library(stringr)
library(tidyft)
library(ggplot2)
library(tidyr)
library(readr)
library(plotly)
library(GGally)

#ensure prints all values in dataframe
options(max.print = 1000000)

#load in the FIFA_players.csv data
fifa <- FIFA_Players

#Part 2.1: Use the string functions and regular expressions to assess a relationship between height and weight among soccer players.  To do this you will need to manipulate the height and weight columns into columns that have numeric values of the height and weight.  
#Create numeric vector based on "height" 

#separate the weight values to make numeric
wt_lbs <- substr(fifa$Weight, 1,3)
wt_lbs <- as.numeric(wt_lbs)

#separate the height values to make numeric 
h <- data.frame(x = fifa$Height)
ht_ft_in <- h %>% separate(x, c("ft", "inch"), sep = "'")
ft <- as.numeric(ht_ft_in$ft)
inch <- as.numeric(ht_ft_in$inch)
ft2in <- ft*12 #convert ft column units from feet to inches 
ht_in <- ft2in+inch #add the two column values together and store in TotalInches

#rebuild the FIFA_Players data.frame with the cleaned up height and weight data
fifa2 <- data.frame(fifa, ht_in, wt_lbs)

fifa2 %>% filter(!is.na(wt_lbs), !is.na(ht_in)) %>% ggplot(mapping = aes(x = ht_in, y = wt_lbs))+geom_point()+geom_smooth()+ggtitle("Relationship Between Weight & Height")+xlab("Height")+ylab("Weight")

#Looking deeper
#What relationship does weight and height have on the overall player skill rating?

#Get summary IQR values for the Overall values to assign future categories
summary(fifa2$Overall)
fifa2_cat <- fifa2 %>% mutate(Skill.Range =  cut(Overall, breaks = c(0, 62, 66, 71, 100),labels = c("Bad", "Okay", "Good", "Great")), .after = Overall) # Breaks are based on the IQR values

#EDA on Players skill range, weight, and height
fifa2_cat %>% filter(!is.na(Skill.Range), !is.na(wt_lbs), !is.na(ht_in)) %>% select(ht_in, wt_lbs, Skill.Range) %>% ggpairs(mapping = aes(fill = Skill.Range))

#break down the skill range category into "Great" and "Not Great" based on IQR data
fifa2_cat2 <- fifa2 %>% mutate(Skill.Range =  cut(Overall, breaks = c(0, 71, 100),labels = c("Not Great", "Great")), .after = Overall) # Breaks are based on the IQR values

#EDA on Players skill range, weight, and height
fifa2_cat2 %>% filter(!is.na(Skill.Range), !is.na(wt_lbs), !is.na(ht_in)) %>% select(ht_in, wt_lbs,Skill.Range) %>% ggpairs(mapping = aes(fill = Skill.Range)) 

#load the two sample t-test with the assumption the "Great" players are significantly taller than "Not Great" players
wt_great <- fifa2_cat2 %>% filter(Skill.Range == "Great") %>% select(wt_lbs) #Define Height variable for great players

wt_notgreat <- fifa2_cat2 %>% filter(Skill.Range == "Not Great") %>% select(wt_lbs) #Define Height variable for not great players

ttest <- t.test(wt_great, wt_notgreat, alternative = "greater", var.equal = FALSE) #test the hypothesis that "Great" players are taller than "Not Great" players. 

ttest


#Part 2.2: Next, assess this relationship between just the LB and LM positions.  (1 slide should do it.)
#EDA on Players weight, and height and LM and LB positions
fifa2 %>% filter(!is.na(wt_lbs), !is.na(ht_in), Position == "LM" | Position == "LB") %>% select(ht_in, wt_lbs, Position) %>% ggpairs(mapping = aes(fill = Position))

#load the two sample t-test for LM and LB weight differences
wt_LM <- fifa2 %>% filter(Position == "LM") %>% select(wt_lbs) #Define weight variable for LM positions

wt_LB <- fifa2 %>% filter(Position == "LB") %>% select(wt_lbs) #Define weight variable for LB positions

ttest_w <- t.test(wt_LB, wt_LM, alternative = "greater", var.equal = FALSE) #test the hypothesis that "Great" players are taller than "Not Great" players. 

ttest_w

#load the two sample t-test for LM and LB height differences
ht_LM <- fifa2 %>% filter(Position == "LM") %>% select(ht_in) #Define weight variable for LM positions

ht_LB <- fifa2 %>% filter(Position == "LB") %>% select(ht_in) #Define weight variable for LB positions

ttest_h <- t.test(ht_LB, ht_LM, alternative = "greater", var.equal = FALSE) #test the hypothesis that "Great" players are taller than "Not Great" players. 

ttest_h

```

Part 4:
Data Munging (30 points): Utilize yob2016.txt for this question. This file is a series of popular children’s names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.

```{r}
#First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.



#Display the summary and structure of df



#Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y’s at the end of the name.  Write an R command to figure out which name it is and display it.




#Upon finding the misspelled name, please remove this particular observation, as the client says it’s redundant.  Save the remaining dataset as an object: y2016 




```



